### 2023-04-18 paper.tex
typeset integer ranges with [1..n]
... instead of [1, \dots, n].

Let $\Theta$ be the parameter (or weight) space, $\theta_t \in \Theta$ $(t \in \natnumzero)$ a sequence of iterates produced by stochastic optimization with $\theta_0$ being the initial value, and $f \colon \Theta \to \Rb$ the generalization loss function.
We may choose the generalization loss to simply be the validation loss, or it may measure performance on a down-stream task.
We assume the generalization loss is evaluated periodically, every $E \in \natnum$ optimization steps, and it is at these points $n \in [0, E, 2E, 3E, \dots]$ where we would like to know how many of the most recent iterates to average to minimize it.
Here and in the following, $t$ and $n$ (with or without subscripts) are assumed to be from $\natnumzero$ and $[0, E, 2E, 3E, \dots]$, respectively, and \emph{loss} always refers to $f$.
Denoting the average of most recent $\Delta$ iterates up to time step $t$ with $\avg(t, \Delta) = \frac{1}{\Delta} \sum_{i=t+1-\Delta}^{t} \theta_i$, we define the optimal averaging length as
\begin{align*}
\OL(t) = \argmin_{\Delta \in [1, \dots, t]} f(\avg(t, \Delta)).
\end{align*}
Our task is to approximate $\OL(n)$ and $\avg(n, \OL(n))$ at all evaluation steps $n$ during optimization.

The trivial algorithm to find $\OL(n)$, which saves all $\theta_i$ and performs a search over $\Delta \in [1, \dots, n]$ to minimize $f(\avg(n, \Delta))$, has prohibitive storage and evaluation cost, proportional to $n$.
Even assuming that $f$ improves monotonically in $\Delta$ up to its optimum, the cost is still proportional to $\OL(n)$.
Our proposed algorithm approximates $\OL(n)$ and $\avg(n, \OL(n))$ with a constant cost.

>>>>

Let $\Theta$ be the parameter (or weight) space, $\theta_t \in \Theta$ $(t \in \natnumzero)$ a sequence of iterates produced by stochastic optimization with $\theta_0$ being the initial value, and $f \colon \Theta \to \Rb$ the generalization loss function.
We may choose the generalization loss to simply be the validation loss, or it may measure performance on a down-stream task.
We assume the generalization loss is evaluated periodically, every $E \in \natnum$ optimization steps, and it is at these points $n \in [0, E, 2E, 3E, \dots]$ where we would like to know how many of the most recent iterates to average to minimize it.
Here and in the following, $t$ and $n$ (with or without subscripts) are assumed to be from $\natnumzero$ and $[0, E, 2E, 3E, \dots]$, respectively, and \emph{loss} always refers to $f$.
Denoting the average of most recent $\Delta$ iterates up to time step $t$ with $\avg(t, \Delta) = \frac{1}{\Delta} \sum_{i=t+1-\Delta}^{t} \theta_i$, we define the optimal averaging length as
\begin{align*}
\OL(t) = \argmin_{\Delta \in [1..t]} f(\avg(t, \Delta)).
\end{align*}
Our task is to approximate $\OL(n)$ and $\avg(n, \OL(n))$ at all evaluation steps $n$ during optimization.

The trivial algorithm to find $\OL(n)$, which saves all $\theta_i$ and performs a search over $\Delta \in [1..n]$ to minimize $f(\avg(n, \Delta))$, has prohibitive storage and evaluation cost, proportional to $n$.
Even assuming that $f$ improves monotonically in $\Delta$ up to its optimum, the cost is still proportional to $\OL(n)$.
Our proposed algorithm approximates $\OL(n)$ and $\avg(n, \OL(n))$ with a constant cost.
